---
title: "R Notebook"
output: html_notebook
---

```{r}
# Import 
library(readxl)
library(dynlm)
library(lmtest)
library(forecast)
library(dplyr)
library(ggplot2)
library(knitr)
```

```{r}
# Read in the data 
rm(list = ls())
setwd('C:/Users/PC/Documents/Tue course/Data Challange 2/JBG050-Data-Challenge-2-Group15')
df = read.csv('data/filter/df_all.csv')
df = data.frame(df)[, -1]
date = df[, 1]

# Get total 
total <- list()
for(row in 1:nrow(df)) {
  total <- append(total, sum(df[row, -1]))  
}
df$total <- total

# Format time-series 
df_total = ts(df['total'], start = c(2010, 12), frequency=12)
```

### 1. Seasonal plot

```{r}
ggseasonplot(df_total, year.labels=TRUE, year.labels.left=TRUE) +
  ylab("Crime") +
  ggtitle("Seasonal plot")
```

```{r}
df_total %>% decompose(type="additive") %>%
  autoplot() + xlab("Year") 
```

X11 decomposition

```{r}
library(seasonal)
df_total %>% seas(x11="") -> fit
autoplot(fit) +
  ggtitle("X11 decomposition of total burglary count")

```

In general, the trend is going down. There are many irregular change between the time of 2019-2021.5 (Covid I believe).

```{r}
par(mfrow=c(2,1))
acfpl <- acf(df_total, plot=TRUE)
pacfpl <- pacf(df_total, plot=TRUE)
```

According to the ACF, the trend-cycle is also 1 year.

### 2. ARIMA model

#### 2.1 Choosing parameters

Let first separate some data for comparison.

```{r}
df <- window(df_total, end=c(2022, 1))
```

```{r}
# seasonal difference 
df %>% diff(lag=12) %>% diff() %>% ggtsdisplay()
```

This is the ACF of the first difference. We observe that lag 1, lag 6, and lag 12 is significant. The lag after lag 1 is mostly insignificant. This suggest AR(1). lag 13 is significant thus suggest order (1) for seasonal part of AR model.

Hyndman-Khandakar algorithm for automatic ARIMA modelling

```{r}
# Using auto-arima(seasonal=TRUE)
fit <- auto.arima(df, seasonal=TRUE)
fit
```

#### 2.2 Check residuals

```{r}
fit %>% residuals() %>% ggtsdisplay()
```

```{r}
checkresiduals(fit)
```

Yep. Residuals is actually white-noise.

#### 2.3 Forecasting

```{r}
fit %>% forecast(h=13) %>% autoplot() +
  autolayer(window(df_total, start=c(2022, 2)), series='Data')
```

```{r}
# Log transformation 
log_df_total <- log(df_total)
log_df <- window(log_df_total, end=c(2022, 1))
log_fit <- auto.arima(log_df, seasonal=TRUE)
log_fit %>% forecast(h=13) %>% autoplot() +
  autolayer(window(log_df_total, start=c(2022, 2)), series='Data')
```

#### 2.4 Add external regressors

For `auto.arima()` to work with external regressors, collect your regressors into a matrix `X`, which you feed into the `xreg` parameter of `auto.arima()`. (Of course, `X` must have the same number of rows as the time series `y` you are modeling.)

For forecasting, you will need the *future* values of your regressors, which you then again feed into the `xreg` parameter of `forecast`.

```{r}
# Before and after covid 
gen_covid <- function(df){
  covid <- (time(df) < 2018.9) | (time(df) > 2021.3)
  in_covid <- ifelse(covid == TRUE, 1, 0)
  
  return(in_covid)
}
# Fit 
fit <- auto.arima(df, seasonal=TRUE, xreg=gen_covid(df))
real <- window(df_total, start=c(2022, 2))
fit %>% forecast(h=13, xreg=gen_covid(real)) %>% autoplot() +
  autolayer(real, series='Data')
```

```{r}
summary(fit)
```

```{r}
# Weather 
setwd('C:/Users/PC/Documents/Tue course/Data Challange 2/JBG050-Data-Challenge-2-Group15')
weather <- read.csv('weatherdata/weather_daily.csv')
weather <- weather %>%
  rename(
     'mean_temp' = 'temperature_2m_mean...C.',
    'rain_total' = 'rain_sum..mm.',
  )
```

```{r}
colnames(weather)
```

```{r}
weather <- weather %>% select('time', 'mean_temp', 'rain_total')
```

```{r}
ts_weather <- ts(weather[-1], start=c(2010, 1), frequency=12)
train_weather <- ts_weather %>% window(start=c(2010, 12), end=c(2021, 1))
test_weather <- ts_weather %>% window(start=c(2021, 2), end=c(2022, 2))
```

```{r}
# Fit 
df <- window(df_total, end=c(2021, 1))
fit <- auto.arima(df, seasonal=TRUE, xreg=train_weather)
real <- window(df_total, start=c(2021, 2))
fit %>% forecast(h=13, xreg=test_weather) %>% autoplot() +
  autolayer(real, series='Data')
```

```{r}
summary(fit)
```

```{r}
checkresiduals(fit)
```

Combine weather and covid data

```{r}
# Combine weather data and covid 
train <- cbind(covid=gen_covid(df), mean_temp=train_weather[,1], rain_total=train_weather[, 2])
real <- window(df_total, start=c(2021, 2), end=c(2022, 2))
test <- cbind(covid=gen_covid(real), mean_temp=test_weather[,1], rain_total=test_weather[, 2])
# Fit
fit <- auto.arima(df, seasonal=TRUE, xreg=train)
fit %>% forecast(h=13, xreg=test) %>% autoplot() +
  autolayer(real, series='Data')

```

```{r}
summary(fit)
```

```{r}
# Check residuals
checkresiduals(fit)
```
